On The Difficulty of Computing Non-Commutative Signature Gröbner Bases

### 2022-01-04 article_issac.tex

For $r \in \NN$ we denote
by $\mathscr{F}_r = \left( \KX \otimes \KX \right)^r$
the free $\KX$-bimodule
of rank $r$, with canonical
basis $\varepsilon_1,\dots,\varepsilon_r$
where $\varepsilon_i =(0,\dots,0,1 \otimes 1,0,\dots,0)$ ($1 \otimes 1$ in the $i$-th position) for $i \in \llbracket 1,r \rrbracket.$
Let $\MM \left( \mathscr{F}_r \right)= \left{ a \varepsilon_i b \mid a,b \in \left\langle X \right\rangle, \: i \in \llbracket 1,r\rrbracket \right}$
be the set of module monomials in $\mathscr{F}_r.$ It is a $K$-vector space basis
of $\mathscr{F}_r.$
Let $\leq_\MM$ be a module
ordering on $\MM \left( \mathscr{F}_r \right).$
We recall that it means that
$\leq_\MM$ is a weel-ordering
compatible with 
scalar multiplication, that
is 
$\mu \leq_\MM \mu'$ implies
$a \mu b \leq_\MM a \mu' b$
for all $\mu,\mu' \in \MM \left( \mathscr{F}_r \right)$
and $a, b \in \left\langle X \right\rangle.$

>>>>

For $r \in \NN$ we denote
by $\mathscr{F}_r = \left( \KX \otimes \KX \right)^r$
the free $\KX$-bimodule
of rank $r$, with canonical
basis $\varepsilon_1,\dots,\varepsilon_r$
where $\varepsilon_i =(0,\dots,0,1 \otimes 1,0,\dots,0)$ ($1 \otimes 1$ in the $i$-th position) for $i \in \llbracket 1,r \rrbracket.$
Let $\MM \left( \mathscr{F}_r \right)= \left\lbrace a \varepsilon_i b \mid a,b \in \left\langle X \right\rangle, \: i \in \llbracket 1,r\rrbracket \right\rbrace$
be the set of module monomials in $\mathscr{F}_r.$ It is a $K$-vector space basis
of $\mathscr{F}_r.$
Let $\leq_\MM$ be a module
ordering on $\MM \left( \mathscr{F}_r \right).$
We recall that it means that
$\leq_\MM$ is a well-ordering
compatible with 
scalar multiplication, that
is 
$\mu \leq_\MM \mu'$ implies
$a \mu b \leq_\MM a \mu' b$
for all $\mu,\mu' \in \MM \left( \mathscr{F}_r \right)$
and $a, b \in \left\langle X \right\rangle.$
We assume that 
$\leq$ and $\leq_\MM$
are compatible
(\textit{i.e.} for any
$a,b \in \left\langle X \right\rangle$, $i \in \llbracket 1,r \rrbracket$,
$a \leq b \Leftrightarrow a \varepsilon_i \leq_\MM b \varepsilon_i \Leftrightarrow \varepsilon_i a \leq_\MM \varepsilon_i b$).
We also assume that 
$\leq_\MM$ is fair (for any
$\mu \in \MM (\mathscr{F}_r)$,
the set $\left\lbrace \mu' \in \MM (\mathscr{F}_r) \mid \mu' \leq_\MM \mu \right\rbrace$ is finite).

If $\alpha \in \mathscr{F}_r$, it can be uniquely
written in the basis 
$\MM \left( \mathscr{F}_r \right) $
and the biggest monomial
involved is its signature
$\mathfrak{s}(\alpha) \in
\mathscr{F}_r$.

If $M \subset \mathscr{F}_r$
is a $\KX$-submodule,
then $G \subset M$ is called
a Gröbner basis of $M$ if
\[\mathfrak{s}(M)= \left\lbrace a \mathfrak{s}(\gamma) b \mid a,b \in \left\langle X \right\rangle, \gamma \in G \right\rbrace. \]

Let $f_1,\dots,f_r \in \KX$
be some polynomials
generating the ideal $I=(f_1,\dots,f_r).$
They define a 
$\KX$-module homomorphism:

\[
\begin{array}{ccccc}
&& \mathscr{F}_r & \rightarrow & \KX \\
\overline{\cdot}& : &\alpha = \sum_i c_i a_i \varepsilon_{j_i} b_i&\mapsto& \overline{\alpha}=\sum_i c_i a_i f_{j_i} b_i. \\
\end{array}\]

Elements $(f,\alpha) \in \KX \times \mathscr{F}_r$
such that $f=\overline{\alpha}$
are called signature polynomials.
We denote by $f^{[\alpha]}$
such a pair. 
For such a pair, necessarily,
$f \in I.$
We denote by $I^{[\Sigma ]}=\left\lbrace f^{[\alpha]} \mid f=\overline{\alpha} \right\rbrace \subset I \times \mathscr{F}_r$
the set of all signature
polynomials.

For $f^{[\alpha]},g^{[\beta]},
h^{[\gamma]} \in I^{\Sigma},$
with $\alpha, g \neq 0,$
we say that 
$f^{[\alpha]}$ $\mathfrak{s}$-reduces to
$g^{[\beta]}$ by
$h^{[\gamma]}$
if there exist $a,b \in 
\left\langle X \right\rangle$
such that:
\begin{itemize}
\item $a \LM(h)b \in supp(f),$
\item $\mathfrak{s}(a \gamma b) \leq \mathfrak{s}(\alpha),$ and
\item $g^{[\beta]}=f^{[\alpha]}-\frac{coeff(f,a \LM(h)b}{\LC(g)} a h^{[\gamma]}b.$
\end{itemize}
Reduction by a set $G^{[\Sigma]} \in I^{[\Sigma]}$
is then defined as the 
reflexive, transitive closure
of the reduction by elements
of $G^{[\Sigma]} $.

Finally, $G^{[\Sigma]} $
is called a \textit{signature
Gröbner basis} of $I^{[\Sigma]}$ if all
$f^{[\alpha]} \in I^{[\Sigma]}$ $\mathfrak{s}$-reduce to zero by $G^{[\Sigma]}.$
